Development of an effective model for SrCuO$_2$ using \textit{ab-initio} quantum Monte Carlo density-matrix downfolding

### 2018-09-27 prelim/draft.tex
Prelim draft ready for Lucas!

\documentclass{article}
\usepackage[margin=1.125in]{geometry}
\usepackage{graphicx}
\usepackage{float}
\usepackage{amsfonts}

\begin{document}
\section{Introduction} 
\textbf{Understanding the low energy behavior of the cuprate superconductors has been an open question in theoretical condensed matter physics for over thirty years.}
The cuprate materials all share a similar structure of repeating layers of square lattice CuO$_2$ planes and charge reservoir blocks, of which carriers on the prior exhibit the most interesting collective behavior. 
At zero doping the cuprates are antiferromagnetic charge-transfer insulators \cite{RevModPhys.78.17}. 
The theoretical challenge arises away from zero doping where the poorly understood superconducting, pseudo-gap, and strange metal phases appear \cite{RevModPhys.84.1383}. 
\ref{fig1} shows the phase diagram of hole- and electron-doped cuprate materials with different structures. 
The superconducting phase is d-wave in character and cannot be described by phonon-induced pairings \cite{PhysRevLett.71.2134,PhysRevLett.69.1431,PhysRevB.43.2778, PhysRevB.47.11314, PhysRevLett.70.1553, RevModPhys.67.515, RevModPhys.77.109, RevModPhys.79.353, RevModPhys.75.473}. 
In the pseudo-gap phase the d-wave gap from the superconducting state continues to partially exist, so that parts of the Fermi surface are gapped above the superconducting transition temperature \cite{ROSSATMIGNOD199186, RevModPhys.77.721, PhysRevLett.75.4114, PhysRevB.61.9752}.
The strange metal phase in the hole doped cuprates is metallic but has a "strange" resistivity which does not depend quadratically on temperature like normal Fermi liquids \cite{PhysRevLett.69.2975}, but does have T-quadratic resitivity for the electron-doped materials \cite{1989PhyC..161..415T}.

@@

\begin{figure}[H]
\centering
\includegraphics[width=0.9\textwidth]{Figures/R2-pareto.pdf}
\caption{\label{fig4} Results of the \textit{ad hoc} fitting procedure just using the DFT eigenvalues and total energies for three different dopings.}
\end{figure}

\begin{figure}[H]
\centering
\includegraphics[width=0.9\textwidth]{Figures/R3-iao_basis.pdf}
\caption{\label{fig5} Relaxation of the IAOs for the spin majority and minority channel under doping for the eighth doped flip state. The majority spin IAO can be used as a basis for the fixed spin moments and the minority spin IAO for the itinerant electrons.}
\end{figure}
\pagebreak

@@

\begin{figure}[H]
\centering
\includegraphics[width=0.9\textwidth]{Figures/P2-Ni_Cu.pdf}
\caption{\label{fig6} The primitive unit cells of the isostructural materials LNCO \cite{PhysRevB.92.205114} and LSCO \cite{Scalapino2012}, the prior which is not super conducting all the way to temperatures of 30mK.}
\end{figure}

>>>>

\documentclass{article}
\usepackage[margin=1.125in]{geometry}
\usepackage{graphicx}
\usepackage{float}
\usepackage{amsfonts}
\usepackage{titling}
\title{Development of an effective model for SrCuO$_2$ using \textit{ab-initio} quantum Monte Carlo density-matrix downfolding}
\author{Shivesh Pathak, Advisor: Lucas Wagner}
\date{TBD}

\begin{document}
\maketitle
\begin{abstract}
\normalsize
Understanding the low energy behavior of the cuprate superconductors has been an open question in theoretical condensed matter physics for over thirty years. Various model Hamiltonians have been proposed and tested to describe different phases under variation in doping and temperature, but the development and validation of these models have not been systematic. We propose that a density matrix downfolding approach \cite{10.3389/fphy.2018.00043, doi:10.1063/1.4927664} using \textit{ab-initio} quantum Monte Carlo can generate and validate an accurate low-energy model for the electron-doped cuprate SrCuO$_2$ at zero temperature and dopings of x=0, 0.125, 0.25 in a systematic way beginning directly from the \textit{ab-initio} Hamiltonian. By conducting a similar downfolding for the isostructural non-superconducting nickelate SrNiO$_2$, and comparing to the effective model for SrCuO$_2$, we hope to develop a better understanding of the mechanisms for superconductivity present in the cuprates. 
\end{abstract}
\pagebreak

\section{Introduction} 
\textbf{Understanding the low energy behavior of the cuprate superconductors has been an open question in theoretical condensed matter physics for over thirty years.}
The cuprate materials all share a similar structure of repeating layers of square lattice CuO$_2$ planes and charge reservoir blocks, of which carriers on the prior exhibit the most interesting collective behavior. 
At zero doping the cuprates are antiferromagnetic charge-transfer insulators \cite{RevModPhys.78.17}. 
The theoretical challenge arises away from zero doping where the poorly understood superconducting, pseudo-gap, and strange metal phases appear \cite{RevModPhys.84.1383}. 
\ref{fig1} shows the phase diagram of hole- and electron-doped cuprate materials with different structures. 
The superconducting phase is d-wave in character and cannot be described by phonon-induced pairings \cite{PhysRevLett.71.2134,PhysRevLett.69.1431,PhysRevB.43.2778, PhysRevB.47.11314, PhysRevLett.70.1553, RevModPhys.67.515, RevModPhys.77.109, RevModPhys.79.353, RevModPhys.75.473}. 
In the pseudo-gap phase the d-wave gap from the superconducting state continues to partially exist, so that parts of the Fermi surface are gapped above the superconducting transition temperature \cite{ROSSATMIGNOD199186, RevModPhys.77.721, PhysRevLett.75.4114, PhysRevB.61.9752}.
The strange metal phase in the hole doped cuprates is metallic but has a "strange" resistivity which does not depend quadratically on temperature like normal Fermi liquids \cite{PhysRevLett.69.2975}, but does have T-quadratic resitivity for the electron-doped materials \cite{1989PhyC..161..415T}.

@@

\begin{figure}[H]
\centering
\includegraphics[width=0.7\textwidth]{Figures/R2-pareto.pdf}
\caption{\label{fig4} Results of the \textit{ad hoc} fitting procedure just using the DFT eigenvalues and total energies for three different dopings.}
\end{figure}

\begin{figure}[H]
\centering
\includegraphics[width=0.7\textwidth]{Figures/R3-iao_basis.pdf}
\caption{\label{fig5} Relaxation of the IAOs for the spin majority and minority channel under doping for the eighth doped flip state. The majority spin IAO can be used as a basis for the fixed spin moments and the minority spin IAO for the itinerant electrons.}
\end{figure}
\pagebreak

@@

\textbf{An approximate timeline for accomplishing each goal follows.} Generation of low-energy states and a complete descriptor space for the cuprate calculation will take the most time in the model fitting process as it requires multiple revisions and will require about two semesters. Studying finite size effects and implementing a similar downfolding to the nickelate material should take a sum total of one semester. Finite size corrections are time-gated only by computational hours (easier to come by than human hours) and the nickelate downfolding should be much easier given the work already done on the isostructural cuprate. Therefore we expect the proposed work to take another three semesters to complete.

\begin{figure}[H]
\centering
\includegraphics[width=0.7\textwidth]{Figures/P2-Ni_Cu.pdf}
\caption{\label{fig6} The primitive unit cells of the isostructural materials LNCO \cite{PhysRevB.92.205114} and LSCO \cite{Scalapino2012}, the prior which is not super conducting all the way to temperatures of 30mK.}
\end{figure}
